### 2019-10-22 pseudo.tex

\renewcommand\thesubfigure{\alph{subfigure}}
\begin{figure*}[ht!]
\centering
\begin{minipage}{.45\linewidth}
\centering
\resizebox{0.925\linewidth}{!}{
\begin{subfigure}[t]{\linewidth}
\inputminted[numbersep=5pt, linenos=true, fontsize=\small]{c}{pseudocode/overall.py}

\caption{The overall framework of \tool}
\label{fig:general_framework}    
\end{subfigure}}\\[0.3em]
\resizebox{0.925\linewidth}{!}{
\begin{subfigure}[t]{\linewidth}
\inputminted[numbersep=5pt, linenos=true, fontsize=\small]{python}{pseudocode/get_cov.py}

\caption{Method to obtain edge coverage}
\label{fig:coverage}    
\end{subfigure}}%~
\end{minipage}\hspace{0.1em}%
\begin{minipage}{.5\linewidth}
\centering
\resizebox{0.925\linewidth}{!}{
\begin{subfigure}{\linewidth}
\inputminted[numbersep=5pt, linenos=true, fontsize=\small]{python}{pseudocode/seed_select.py}
\vspace{-2em}
\caption{Method to perform seed selection with importance sampling.}
\label{fig:seed_selection}    
\end{subfigure}}\\[-0.2em]
\resizebox{0.925\linewidth}{!}{
\begin{subfigure}{\linewidth}
\inputminted[numbersep=5pt, linenos=true, fontsize=\small]{python}{pseudocode/train_adaptive.py}
\vspace{-2em}
\caption{Method to compute adaptive loss for training neural network.}
\label{fig:train_adaptive}    
\end{subfigure}}\\[-0.2em]
\resizebox{0.925\linewidth}{!}{        
\begin{subfigure}{\linewidth}
\inputminted[numbersep=5pt, linenos=true, fontsize=\small]{python}{pseudocode/mutate.py}
\vspace{-2em}
\caption{Method to generate new inputs with gradient ascent.}
\label{fig:mutation}    
\end{subfigure}}
\end{minipage}\\[0.3em]
\caption{{\small Pseudocode of \tool.}}
\label{fig:pseudo}
\end{figure*}